---
title:  "TCR Analysis Report<br>Conroy | Naidoo"
date:   "`r paste0('Date: ', format(Sys.time(), '%Y-%m-%d %X %Z'))`"
author: "Prepared by James Robert White, PhD (Resphera Biosciences)" 
output:
  html_document:
    theme: default
    highlight: default
    toc: true
    toc_depth: 3
    collapse: subsubsection
    number_sections: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---
<style type="text/css">
.list-group-item.active {
  color: white;
  background-color: #1A2188;
  border: none;
}
.list-group-item:hover,
.list-group-item.active:hover {
    color: white;
    background-color: #1A2188;
}
.title{
  color: #1A2188;
}
.author{
  color: #3A5CE9;
}
.date{
  color: #4a4e69;
}
h1 {
  font-size: 28px;
}
h2 {
  font-size: 26px;
}
h3 {
  font-size: 24px;
}
h4 {
  font-size: 20px;
}
</style>


```{r setup, echo=FALSE, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo=FALSE)
library(caret)
library(DescTools)
library(dplyr)
library(DT)
library(table1)
library(egg)
library(ggplot2)
library(ggpubr)
library(ggrepel)
library(grid)
library(pROC)
library(ROCR)
library(randomForest)
library(reshape2)
library(stringr) 
library(scales)

# ---
# to run:
# setwd("~/Desktop/jhmi-neoadj-mbiom-2024/code")
# library(rmarkdown)
# render("A02.0-tcr-eval.Rmd")

# --
# manual options
setwd("~/Desktop/jhmi-neoadj-mbiom-2024/code")
# ---
# name output directory
script_prefix = "A02.0-tcr-eval"
# ---
# create output directory, overwrite if needed
output_dir = paste("../analysis/", script_prefix, sep="")
unlink(output_dir, recursive=TRUE)
dir.create(output_dir)

# ---
# store results
my_fig_list      = list()
my_stat_list     = list()
my_table_list    = list()
my_by_level_list = list()

files  = c("jhmi-neoadj-mbiom-2024.alpha-diversity.txt",
           "jhmi-neoadj-mbiom-2024.phylum.txt",
           "jhmi-neoadj-mbiom-2024.class.txt",
           "jhmi-neoadj-mbiom-2024.order.txt",
           "jhmi-neoadj-mbiom-2024.family.txt",
           "jhmi-neoadj-mbiom-2024.genus.txt",
           "jhmi-neoadj-mbiom-2024.picrust.txt",
           "jhmi-neoadj-mbiom-2024.picrust2.pw.txt")

# --
# data files

flx_table_1 = c()
flx_table_2 = c()

for (fi in files){
  lvl = gsub(".pw", "", gsub("-", " ", gsub(".avgGTE0.01Prct", "", gsub(".txt", "", gsub("jhmi-neoadj-mbiom-2024.", "", fi)))))
  print(lvl)


  fi_df = read.table(paste0("../analysis/A02.0-add-metadata/", fi), header=TRUE, sep="\t", check.names=FALSE)

  # drop non tcr samples
  fi_df = fi_df[fi_df$TCRseq==1,]

  # remove all zero columns
  keep_set = colnames(fi_df[,26:ncol(fi_df)])[colSums(fi_df[,26:ncol(fi_df)]) > 0]
  fi_df    = fi_df[,c(colnames(fi_df)[1:25], keep_set)]
  # drop timepoints before 90 days
  fi_df    = fi_df[fi_df$`days from start` <= 90,]
  # fi_df[,2] = cut(fi_df$`days from start`, breaks=c(-10, 90, Inf))
  # colnames(fi_df)[2] = "time"
  
  fi_features = colnames(fi_df)[26:ncol(fi_df)] 
  targets = c("total productive reads", "max productive freq", "total clonotypes", "clonality")
  
  for (j in targets){
    fi_df[,j] = as.numeric(as.character(fi_df[,j]))
  }  
  for (j in fi_features){
    fi_df[,j] = as.numeric(as.character(fi_df[,j]))
  }  

  stats_df = c()

  for (fif in fi_features){
    # evaluate each target column for this feature
    # get spearman and pearson correlation
    # with p-values
    
    for (trg in targets){
      fif_row = c(lvl, fif)
      sp.cor = cor.test(fi_df[,trg], fi_df[,fif], method="spearman")
      pr.cor = cor.test(fi_df[,trg], fi_df[,fif], method="pearson")
      fif_row = c(fif_row, trg, sp.cor$estimate, sp.cor$p.value, 
                           pr.cor$estimate, pr.cor$p.value)

      stats_df = rbind(stats_df, fif_row)
  
    }
  }

  colnames(stats_df) = c("level", "feature", "tcr measure", "spearman est", "sp pval", "pearson est", "pr pval") 
  stats_df = data.frame(stats_df, check.names=FALSE)
  for (j in 4:7){
    stats_df[,j] = round(as.numeric(stats_df[,j]), digits=4)
    stats_df[is.na(stats_df[,j]),j] = "n/a"
  }
   
  my_by_level_list[[paste0(lvl, " stats")]]  <- datatable(data = stats_df, 
                                                        class      = 'cell-border stripe', 
                                                        rownames   = F, 
                                                        extensions = "Buttons", 
                                                        filter     = "top",
                                                        options    = list(paging    = TRUE,
                                                                          scrollX   = TRUE,
                                                                          searching = TRUE,  
                                                                          dom = 'Bfrtip',
                                                                          buttons = c('copy', 'csv', 'excel'),
                                                                          pageLength = 10)) 


  if (grepl("alpha", lvl)){ 

    flx_table_1 <- datatable(data = unique(fi_df[,c("pid", "histology", "stage", "sex", "age", "MPR", "smk", "tx_type", "percent_RVT", targets)]), 
                            class      = 'cell-border stripe', 
                            rownames   = F, 
                             filter     = "top",
                            extensions = "Buttons", 
                            options    = list(paging    = TRUE,
                                              scrollX   = TRUE,
                                              searching = TRUE,  
                                              dom = 'Bfrtip',
                                              buttons = c('copy', 'csv', 'excel'),
                                              pageLength = 20)) 

     flx_table_2 <- datatable(data = unique(fi_df[,c("pid", "days from start", "histology", "stage", "sex", "age", "MPR", "smk", "tx_type", "percent_RVT", targets)]), 
                            class      = 'cell-border stripe', 
                            rownames   = F, 
                             filter     = "top",
                            extensions = "Buttons", 
                            options    = list(paging    = TRUE,
                                              scrollX   = TRUE,
                                              searching = TRUE,  
                                              dom = 'Bfrtip',
                                              buttons = c('copy', 'csv', 'excel'),
                                              pageLength = 20)) 
    
  }


}





```

# Purpose
The purpose of this report is to provide a preliminary data description and 
analysis of TCR feature values related to microbiome profiles (taxonomic and 
functional) associated with the neoadjuvant cohort.

# Scope
This analysis covers the following target TCR variables:

* Total Productive Reads
* Max Productive Freq
* Total Clonotypes
* Clonality

and their associations with the following microbiome levels:

* Alpha diversity
* Phylum
* Class
* Order
* Family
* Genus
* PICRUST v1
* PICRUST v2


# Source Files
Primary resources used in this analysis:

* `Sequencing Data\2023 Lung neoadj-meso samples\TCRseq` Sears lab server
* `jhmi-neoadj-mbiom-2024` Resphera managed microbiome analysis repository

# Dataset for Analysis
Below is a summary of the subject IDs and their metadata. Samples without TCR measures 
and timepoints longer than 90 days have been removed.

```{r echo=FALSE, warning=FALSE}
  flx_table_1
```

Below is a summary of the timepoints per subject ID analyzed.

```{r echo=FALSE, warning=FALSE}
  flx_table_2
```


# Results

## Results by Level

Below for each level we display the primary statistical analysis results
comparing TCR measures and microbiome features using 

1. Spearman correlation
2. Pearson correlation

```{r echo=FALSE, warning=FALSE, fig.width=8.5, fig.height=7.5, fig.align='center', results="asis"}
  for (analyte in names(my_by_level_list)){ 
    cat("\n") 
    cat("\n") 


    cat("###", str_to_title(analyte), "\n\n")
  
    print(htmltools::tagList(my_by_level_list[[analyte]]))

    cat("<br>\n") 
    cat("<br>\n") 
    cat("<br>\n") 

    cat("<br>\n") 
    cat("<br>\n") 
    cat("<br>\n") 

    cat("<br>\n") 
    cat("<br>\n") 
    cat("<br>\n") 

    cat("<br>\n") 
    cat("<br>\n") 
    cat("<br>\n") 

    cat("<br>\n") 
    cat("<br>\n") 
    cat("<br>\n") 

    cat("<br>\n") 
    cat("<br>\n") 
    cat("<br>\n") 

    cat("<br>\n") 
    cat("<br>\n") 
    cat("<br>\n") 

    cat("<br>\n") 
    cat("<br>\n") 
    cat("<br>\n") 


    cat("\n") 
    cat("\n") 
    cat("\n") 
    cat("\n") 
  } 
```


# End of Document
